---
title: "Employee Attrition Prediction"
author: "pradeepkumar, Dan Crouthamel"
date: "8/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r }
library(ggplot2)
library(dplyr)
# Read Data
EmpAttrData <- read.csv("CaseStudy2-data.csv", header = TRUE)
```
# Explorotory Analysis of Data
## Explore corolation among continious explanatory variables
```{r}
###  get correlation matrix and correlation info
library(corrplot)
# Get Only numeric columns
EmpAttrData_Numeric <- EmpAttrData[, sapply(EmpAttrData,is.integer)]
correlations <- cor(EmpAttrData_Numeric)
corrplot(correlations, method="circle")
```

## Is there any missing data? , lets check visually
### Missing data will appear in Blue
```{r}
library(Amelia)
library(mlbench)
missmap(EmpAttrData, col=c("blue", "red"), legend=FALSE)
```

## Explore distribution of explanatory variables
```{r}
# Visualizing distribution of explanatory vars
# Histogram
par(mfrow=c(1,4))
for(i in 1:4) {
  boxplot(EmpAttrData_Numeric[,i], main=paste(names(EmpAttrData_Numeric)[i]), xlab=NA)
}
```




## Pair wise correlation matrix
```{r}
## Pair wise scatter plot
pairs(EmpAttrData_Numeric, col=EmpAttrData$Attrition)
```

## plot of age and count Grouped by job satisfaction

```{r}
t1<-table(EmpAttrData$Age, EmpAttrData$JobSatisfaction)
d1 <- data.frame(t1)
names(d1) <- c('Age','Levels','count')
ggplot(d1, aes(d1$Age,d1$count, color=d1$Levels)) + geom_point() + 
  labs(title="Plot of Age vs Count , Group by Satisfiction",x="Age", y = "Count") +
  theme(legend.title = element_blank())
```

## plot of Age vs Jobsatisfaction , grouped by Gender
```{r}
ggplot(EmpAttrData, aes(EmpAttrData$Gender, EmpAttrData$Age, color=EmpAttrData$Gender)) +
aes(EmpAttrData$JobSatisfaction,EmpAttrData$Age) + geom_boxplot() +
  labs(title="Plot of Age vs JobSatisfaction , Group by Gender",x="JobSatisfactionLevels", y = "Age") +
  theme_classic()
```

## Age vs Gender grouped by Job satisfaction

```{r}
EmpAttrData$JobSatisfaction <- as.factor(EmpAttrData$JobSatisfaction)

ggplot(EmpAttrData, aes(EmpAttrData$Gender, EmpAttrData$Age, color=EmpAttrData$Gender)) +
  aes(EmpAttrData$JobSatisfaction,EmpAttrData$Age) + geom_boxplot() +
  labs(title="Plot of Age vs JobSatisfaction , Group by Gender",x="JobSatisfactionLevels", y = "Age") +
  theme_classic() + theme(legend.title = element_blank())
```

## Years at company and attrition
```{r}
t1<-table(EmpAttrData$YearsAtCompany, EmpAttrData$Attrition)
d1 <- data.frame(t1)
names(d1) <- c('YearsAt','Attrition','count')
ggplot(d1, aes(d1$YearsAt,d1$count, color=d1$Attrition)) + geom_point() + 
  labs(title="Plot of YearsAtCompany vs Count , Group by Attrition",x="YearsAtCompany", y = "Count") +
  theme(legend.title = element_blank())
```

## Explore the distribution of explanatory variables for each 
## level of dependent variable ( attrition)
```{r}
## Density divided by attrition
library(caret)
x <- EmpAttrData[,28:34]
y <- EmpAttrData$Attrition
scales <- list(x=list(relation="free"), y=list(relation="free"))
featurePlot(x=x, y=y, plot="density", scales=scales)
```

```{r}
# Overtime
EmpAttrData %>% ggplot() +
  geom_bar(aes(OverTime, fill= Attrition), position = "stack")+
  theme(plot.title = element_text(hjust = 0.5))+
  ggtitle(label = 'The Effect of Overtime on Attrition')

EmpAttrData %>% ggplot(aes(OverTime, fill= Attrition)) +
  geom_bar(aes(y = (..count..)/sum(..count..)), position = "fill")+
  ggtitle(label = 'The Effect of Overtime on Attrition Rate')+
  theme(plot.title = element_text(hjust = 0.5))+
  ylab('Percentages')

#Job Role & Attrition
EmpAttrData %>% ggplot() +
  geom_bar(aes(JobRole, fill= Attrition), position = "stack")+
  scale_fill_manual(values = c("Yes" = "red","No" = "dark green"))+
  ggtitle(label = 'The Effect of Job role on Attrition')+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

EmpAttrData %>%ggplot(aes(JobRole, fill= Attrition)) +
  geom_bar(aes(y = (..count..)/sum(..count..)), position = "fill")+
  ggtitle(label = 'The Effect of Overtime on Attrition Rate')+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
ylab('Percentages')

#Job Role & Overtime
EmpAttrData %>% ggplot() +
  geom_bar(aes(JobRole, fill= OverTime), position = "stack")+
  scale_fill_manual(values = c("Yes" = "red","No" = "dark green"))+
  ggtitle(label = 'The Effect of Job role on Overtime')+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

EmpAttrData %>%ggplot(aes(JobRole, fill= OverTime)) +
  geom_bar(aes(y = (..count..)/sum(..count..)), position = "fill")+
  ggtitle(label = 'The Effect of Overtime on Overtime Rate')+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
ylab('Percentages')

# Job Satisfaction
EmpAttrData %>% ggplot() +
  geom_bar(aes(JobSatisfaction, fill= Attrition), position = "stack")+
  theme(plot.title = element_text(hjust = 0.5))+
  ggtitle(label = 'The Effect of Job Satisfaction on Attrition')

EmpAttrData %>% ggplot(aes(JobSatisfaction, fill= Attrition)) +
  geom_bar(aes(y = (..count..)/sum(..count..)), position = "fill")+
  ggtitle(label = 'The Effect of Job Satisfaction on Attrition Rate')+
  theme(plot.title = element_text(hjust = 0.5))+
  ylab('Percentages')

# Job Involvement
EmpAttrData %>% ggplot() +
  geom_bar(aes(JobInvolvement, fill= Attrition), position = "stack")+
  theme(plot.title = element_text(hjust = 0.5))+
  ggtitle(label = 'The Effect of Job Involvement on Attrition')

EmpAttrData %>% ggplot(aes(JobInvolvement, fill= Attrition)) +
  geom_bar(aes(y = (..count..)/sum(..count..)), position = "fill")+
  ggtitle(label = 'The Effect of Job Involvement on Attrition Rate')+
  theme(plot.title = element_text(hjust = 0.5))+
  ylab('Percentages')



```

# Build a Logistic Model
```{r }
# Average Attrition Rate (16%)
attritionRate <- sum(EmpAttrData$Attrition == "Yes")/nrow(EmpAttrData)

# Remove the column with only one factor, otherwise glm won't work.
# You'll get a contrasts can be applied only to factors with 2 or more levels.
# Since all values are the same, this affects nothing
EmpAttrData$Over18 <- NULL

# Let's do some analysis to see what affects attrition
glmInfo <- glm(EmpAttrData$Attrition ~ ., family = binomial(link="logit"), data = EmpAttrData)
summary(glmInfo)
```

